---
title: "Linkfire Project"
author: "Tyler Chiu"
date: "5/15/2020"
output: 
  html_document:
    toc: true
    toc_float: true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(plotly)
```

# Assumptions <br>
- winback subscribers were members, then began a trial and became members again <br>
- visitor token matches ip or device address <br>
- winbacks are using different email/phone numbers <br>

# Importing Data
```{r}
artists = read.csv("Artists.csv")
events = read.csv("Events.csv")
linkartists = read.csv("LinkArtists.csv")
```
# Data Processing
```{r}
# converting data into propper format
events = events %>% mutate(Timestamp = ymd_hms(events$Timestamp),
                           Year = year(Timestamp),
                           Month = month(Timestamp),
                           Day = day(Timestamp),
                           Hour = hour(Timestamp),
                           Quarter = quarter(Timestamp)
                           )

# modifying column names to match 
colnames(linkartists)[which(names(linkartists)=="X.ArtistId")] ="ArtistId"
colnames(events)[which(names(events)=="Linkid")] ="LinkId"

#combining tables
linkartists = linkartists %>% left_join(artists)

#switching LinkId to lowercase to match events data
linkartists = linkartists %>% mutate(LinkId = tolower(LinkId))  

#joining linkartists to match LinkId with Artist
events = events %>% left_join(linkartists) 
```


# Data Exploration
Looking at most frequent or most popular values in each field
```{r}
# Summary of browsers used in descending order
sort(summary(events$Browser),decreasing=T) 

# Top 5 Cities
head(sort(summary(events$City),decreasing=T))

# Top 5 countries
head(sort(summary(events$Countrycode),decreasing=T))

# Table of Member Types 
sort(summary(events$Convtype),decreasing=T)

# Table of Member Plans
sort(summary(events$Convname),decreasing=T)

# Table of Devices
sort(summary(events$Device),decreasing=T)
```
# Finding New Subscribers and Winback Customers
Method
I wanted to looking at Linkfire visitors who had visited at least twice in order to see whether or not they were new subscribers or winback customers.

I assumed that visitors who were once on trial and then began memberships became new subscribers.

For Winback customers, I looked for users who were members before going on a trial period and then back to being members. I also counted the number of times a single user resubscribed to their streaming service if they switched between trial and membership services multiple times.

```{r}
repeated_id = events$Visitortoken[duplicated(events$Visitortoken)]

# All users that have visited Linkfire more than once
repeated_id = unique(repeated_id) 

events_repeat = events[events$Visitortoken %in% repeated_id,]

# Most frequent users
head(sort(summary(events$Visitortoken),decreasing = T)) 
events_repeat = events_repeat[order(events_repeat$Visitortoken,events_repeat$Timestamp),]

events_repeat$new_sub = rep(0, nrow(events_repeat))
events_repeat$win_back = rep(0,nrow(events_repeat))

new_id = 0
lost = 0
was_member = 0
on_trial = 0

for(i in 1:nrow(events_repeat)) {
  # Checks if looking at new user
  if(new_id == 1 && events_repeat$Visitortoken[i-1]!=events_repeat$Visitortoken[i]) {
    new_id = 0
    lost = 0
    was_member = 0
    on_trial = 0
  }
  # Assigns values if this is a user's first visit
  if(new_id == 0) { 
    new_id=1
    if(events_repeat$Convtype[i]=="trial") {on_trial=1; 
    } else if(events_repeat$Convtype[i]=="membership") {was_member=1; }
    next
  } 
  # Checks to see if user becomes new subscriber or winback
  if(new_id == 1) { 
    if(events_repeat$Convtype[i]=="trial" && on_trial==1 && lost == 0) { 
    } else if(events_repeat$Convtype[i] == "membership" && was_member == 1 && lost == 0) { on_trial = 0; was_member=1;
    } else if(events_repeat$Convtype[i] == "trial" && was_member == 1) {lost=1; on_trial= 0;
    } else if(events_repeat$Convtype[i] == "membership" && lost == 1 && on_trial==0 && was_member==1) {events_repeat$win_back[i]=1; lost=0;
    } else if(events_repeat$Convtype[i] == "membership" && on_trial==1 && lost == 0) {events_repeat$new_sub[i]=1;on_trial=0;was_member=1;}
  }
}
```

# Number of New Subscribers and Winbacks
```{r}
# Number new subscribed members
sum(events_repeat$new_sub)
# Instances of customers returning to their subscriptions   
sum(events_repeat$win_back) 
```

# Ratio of New Subscribers and Winbacks to Total Amount of Linkfire Visitors
```{r}
# Number of unique visitors total
length(unique(events$Visitortoken)) 

# 17% of Linkfire visitors who become new subscribers
2480/14507 * 100 

# .05% of Linkfire visitors who resubscribe 
8/14507 * 100 
```
# Total Number of Unique Linkfire visitors in 2019 and 2020
```{r}
# Number of unique linkfire visitors in 2019
length(unique(events$Visitortoken[which(events$Year==2019,)]))

# Number of unique linkfire visitors in 2020
length(unique(events$Visitortoken[which(events$Year==2020,)]))

# Percent change in Linkfire visitors from 2019 to 2020
((9811-6382)/6382)*100
```


# Tables
Comparing different fields by the percentage of new subscribers or Winbacks
```{r}
# City with highest percentage of new subscribers 
city_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$City, data = events_repeat, FUN = sum)
city_new_sub = city_new_sub %>% arrange(desc(city_new_sub[,2]))
head(city_new_sub)

# City with highest percentage of winbacks
city_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$City, data = events_repeat, FUN = sum)
city_win_back = city_win_back %>% arrange(desc(city_win_back[,2]))
head(city_win_back)

# Countries with highest percentage of new subscribers
country_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Countrycode, data = events_repeat, FUN = sum)
country_new_sub = country_new_sub %>% arrange(desc(country_new_sub[,2]))
head(country_new_sub) 

# Countries with highest percentage of winbacks
country_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Countrycode, data = events_repeat, FUN = sum)
country_win_back = country_win_back %>% arrange(desc(country_win_back[,2]))
head(country_win_back) 

# Subscription Plans and percentage of new subscribers
plan_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Convname,FUN = sum)
plan_new_sub = plan_new_sub %>% arrange(desc(plan_new_sub[,2]))
head(plan_new_sub) 

# Subscription Plans and percentage of winbacks
plan_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Convname,FUN = sum)
plan_win_back = plan_win_back %>% arrange(desc(plan_win_back[,2]))
head(plan_win_back)

# Devices and percentage of new customers
device_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Device,FUN = sum)
device_new_sub = device_new_sub %>% arrange(desc(device_new_sub[,2]))
device_new_sub 

# Devices and percentage of winback
device_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Device,FUN = sum)
device_win_back = device_win_back %>% arrange(desc(device_win_back[,2]))
device_win_back 

# Years and percentage of new subscribers
year_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Year,FUN = sum)
year_new_sub = year_new_sub %>% arrange(desc(year_new_sub[,2]))
year_new_sub 

# Years and percentage of winback
year_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Year,FUN = sum)
year_win_back = year_win_back %>% arrange(desc(year_win_back[,2]))
year_win_back 


# Months and percentage of new subscribers
month_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Month,FUN = sum)
month_new_sub = month_new_sub %>% arrange((month_new_sub[,1]))
month_new_sub

# Months and percentage of winback
month_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Month,FUN = sum)
month_win_back = month_win_back %>% arrange((month_win_back[,1]))
month_win_back 

# Artists with the highest percentage of new subscribers
artist_new_sub = aggregate(events_repeat$new_sub/2480*100~events_repeat$Name,FUN = sum)
artist_new_sub = artist_new_sub %>% arrange(desc(artist_new_sub[,2]))
artist_new_sub[1:10,]

# Artists with the highest percentage of winback
artist_win_back = aggregate(events_repeat$win_back/8*100~events_repeat$Name,FUN = sum)
artist_win_back = artist_win_back %>% arrange(desc(artist_win_back[,2]))
artist_win_back[1:10,] 
```

# Percent Change in New Subscribers and Winbacks Who Used Linkfire From 2019 to 2020
```{r}
year_new_sub = aggregate(events_repeat$new_sub~events_repeat$Year,FUN = sum)

# Percentage change in number of new subscribers who used Linkfire from 2019 to 2020
(1937-543)/543*100

year_win_back = aggregate(events_repeat$win_back~events_repeat$Year,FUN = sum)
# Percentage change in number of winbacks who used Linkfire from 2019 to 2020
(6-2)/2*100
```


# Interactive Visualizations
(hover over the bars to reveal more info)
```{r}
g = ggplot(device_new_sub, aes(x=device_new_sub[,1],y=device_new_sub[,2])) + geom_bar(stat="identity") + xlab("Devices") + ylab("Number of New Subscribers") +ggtitle("Devices used by New Subscribers") 
ggplotly(g) # people who are new subscribers are on phones

g = ggplot(device_win_back, aes(x=device_win_back[,1],y=device_win_back[,2])) + geom_bar(stat="identity") + xlab("Devices") + ylab("Number of Winbacks") +ggtitle("Devices used by Winbacks")
ggplotly(g) # people who are won back are on desktops

g = ggplot(month_new_sub, aes(x=month_new_sub[,1],y=month_new_sub[,2])) + geom_bar(stat="identity") + xlab("Months") + ylab("Number of New Subscribers") +ggtitle("Monthly count of New Subscribers")
ggplotly(g) # there is an increase 

g = ggplot(month_win_back, aes(x=month_win_back[,1],y=month_win_back[,2])) + geom_bar(stat="identity") + xlab("Months") + ylab("Number of Winbacks") +ggtitle("Monthly count of Winbacks")
ggplotly(g)

g = ggplot(year_new_sub, aes(x=year_new_sub[,1],y=year_new_sub[,2])) + geom_bar(stat="identity") + xlab("Year") + ylab("Number of New Subscribers") +ggtitle("Yearly count of New Subscribers")
ggplotly(g)

g = ggplot(year_win_back, aes(x=year_win_back[,1],y=year_win_back[,2])) + geom_bar(stat="identity") + xlab("Year") + ylab("Number of Winbacks") +ggtitle("Yearly count of Winbacks")
ggplotly(g)

g = ggplot(artist_new_sub[1:10,], aes(x=artist_new_sub[1:10,1],y=artist_new_sub[1:10,2])) + geom_bar(stat="identity") + xlab("Artists") + ylab("Number of New Subscribers") +ggtitle("Artists with most New Subscribers")
ggplotly(g)

g = ggplot(artist_win_back[1:3,], aes(x=artist_win_back[1:3,1],y=artist_win_back[1:3,2])) + geom_bar(stat="identity") + xlab("Artists") + ylab("Number of Winbacks") +ggtitle("Artists with most Winbacks")
ggplotly(g)
```





